---
title: "RAC_All_Tract"
author: "David J Degnan"
date: "5/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warnings = FALSE, message = FALSE)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(trelliscopejs)
library(plotly)
library(DT)
```

## Brief Description

RAC - Resident Area Characteristic Data - totaled by HOME census block

Both federal and non-federal datasets were combined. 

```{r}
RAC_NoFed <- fread("../../og_data/Job_counts/rac_all_tract_minus_fed.csv")
RAC_Fed <- fread("../../og_data/Job_counts/rac_fed_tract.csv")
RAC <- rbind(RAC_NoFed, RAC_Fed)
```

## Subset and Roll Up

Subset to the 2015-2018 years and roll up to the county level.

```{r}
RAC_County <- RAC %>%
  subset(year %in% 2015:2018) %>%
  select(cty, year, colnames(RAC)[8:48]) %>%
  group_by(cty, year) %>%
  summarise(across(everything(), sum)) %>%
  ungroup() %>%
  select(-year) %>%
  group_by(cty) %>%
  summarise(across(everything(), mean)) %>%
  ungroup() %>%
  rename(FIPS = cty)
```

## Get Proportions

*Total:* C000 Num Total number of jobs

*Age*

* CA01 Num Number of jobs for workers age 29 or younger

* CA02 Num Number of jobs for workers age 30 to 54

* CA03 Num Number of jobs for workers age 55 or older

*Earnings*

* CE01 Num Number of jobs with earnings 1250/month or less

* CE02 Num Number of jobs with earnings 1251/month to 3333/month

* CE03 Num Number of jobs with earnings greater than 3333/month

*Industry*

* CNS01 Num Number of jobs in NAICS sector 11 (Agriculture, Forestry, Fishing and Hunting)

* CNS02 Num Number of jobs in NAICS sector 21 (Mining, Quarrying, and Oil and Gas Extraction)

* CNS03 Num Number of jobs in NAICS sector 22 (Utilities)

* CNS04 Num Number of jobs in NAICS sector 23 (Construction)

* CNS05 Num Number of jobs in NAICS sector 31-33 (Manufacturing)

* CNS06 Num Number of jobs in NAICS sector 42 (Wholesale Trade)

* CNS07 Num Number of jobs in NAICS sector 44-45 (Retail Trade)

* CNS08 Num Number of jobs in NAICS sector 48-49 (Transportation and Warehousing)

* CNS09 Num Number of jobs in NAICS sector 51 (Information)

* CNS10 Num Number of jobs in NAICS sector 52 (Finance and Insurance)

* CNS11 Num Number of jobs in NAICS sector 53 (Real Estate and Rental and Leasing)

* CNS12 Num Number of jobs in NAICS sector 54 (Professional, Scientific, and Technical Services)

* CNS13 Num Number of jobs in NAICS sector 55 (Management of Companies and Enterprises)

* CNS14 Num Number of jobs in NAICS sector 56 (Administrative and Support and Waste Management and * 
Remediation Services)

* CNS15 Num Number of jobs in NAICS sector 61 (Educational Services)

* CNS16 Num Number of jobs in NAICS sector 62 (Health Care and Social Assistance)

* CNS17 Num Number of jobs in NAICS sector 71 (Arts, Entertainment, and Recreation)

* CNS18 Num Number of jobs in NAICS sector 72 (Accommodation and Food Services)

* CNS19 Num Number of jobs in NAICS sector 81 (Other Services [except Public Administration])

* CNS20 Num Number of jobs in NAICS sector 92 (Public Administration)

*Race* 

* CR01 Num Number of jobs for workers with Race: White, Alone (split out Hispanic)

* CR02 Num Number of jobs for workers with Race: Black or African American Alone

* CR03 Num Number of jobs for workers with Race: American Indian or Alaska Native Alone

* CR04 Num Number of jobs for workers with Race: Asian Alone

* CR05 Num Number of jobs for workers with Race: Native Hawaiian or Other Pacific Islander Alone

* CR07 Num Number of jobs for workers with Race: Two or More Race Groups

*Ethnicity*

* CT01 Num Number of jobs for workers with Ethnicity: Not Hispanic or Latino

* CT02 Num Number of jobs for workers with Ethnicity: Hispanic or Latino

*Education*

* CD01 Num Number of jobs for workers with Educational Attainment: Less than high school

* CD02 Num Number of jobs for workers with Educational Attainment: High school or equivalent, no college

* CD03 Num Number of jobs for workers with Educational Attainment: Some college or Associate degree

* CD04 Num Number of jobs for workers with Educational Attainment: Bachelor's degree or advanced degree

*Sex*

* CS01 Num Number of jobs for workers with Sex Male

* CS02 Num Number of jobs for workers with Sex Female


```{r}
# Data descriptors
Descriptors <- data.table(
  Column = colnames(RAC_County)[3:ncol(RAC_County)],
  Description = c("Age 14 to 29", "Age 30 to 54", "Age 55 and Older",
                  "Earning 1250 or less", "Earning 1250 to 3333", "Earning 3333 or more",
                  "Agriculture", "Mining", "Utilities", "Construction",
                  "Manufacturing", "Wholesale", "Retail", "Transportation",
                  "Information", "Insurance", "Real Estate", "Science",
                  "Management", "Waste Management", "Education", "Health Care",
                  "Entertainment", "Food Services", "Other Services", "Public Administration",
                  "White", "Black", "Native", "Asian", "Pacific Islander", "Multiracial",
                  "Not Hispanic", "Hispanic", "Less than high school", 
                  "High school", "Some college or Associates", "Bachelors or higher",
                  "Male", "Female")
)

# Calculate each group over total 
Props <- do.call(cbind, lapply(colnames(RAC_County)[3:ncol(RAC_County)], function(name) {
  RAC_County[[name]] / RAC_County$c000
})) %>% data.table()
colnames(Props) <- Descriptors$Description
Props <- data.table(RAC_County[,c(1:2)], Props)
colnames(Props)[2] <- "Average of Annual Total"
fwrite(Props, "Job_Proportions.csv")
datatable(Descriptors)
```

## County Sizes

```{r}
RAC_County$c000 %>% summary()
```


## Jobs per...

We have 3142 samples. 

#### Age

```{r}
Props %>%
  select(Descriptors$Description[1:3]) %>% 
  pivot_longer(Descriptors$Description[1:3]) %>%
  rename(Age = name, Proportion = value) %>%
  ggplot(aes(x = Age, y = Proportion, fill = Age)) + geom_boxplot() + theme_bw() + 
  ylim(c(0,1)) + theme(legend.position = "none")
```

```{r}
pairs(Props %>% select(Descriptors$Description[1:3]))
```

#### Earnings

```{r}
Props %>%
  select(Descriptors$Description[4:6]) %>% 
  pivot_longer(Descriptors$Description[4:6]) %>%
  rename(Earnings = name, Proportion = value) %>%
  ggplot(aes(x = Earnings, y = Proportion, fill = Earnings)) + geom_boxplot() + theme_bw() + 
  ylim(c(0,1)) + theme(legend.position = "none")
```

```{r}
pairs(Props %>% select(Descriptors$Description[4:6]))
```

#### Industry 

```{r}
Props %>%
  select(Descriptors$Description[7:26]) %>% 
  pivot_longer(Descriptors$Description[7:26]) %>%
  rename(Industry = name, Proportion = value) %>%
  ggplot(aes(x = Industry, y = Proportion)) + geom_boxplot() + theme_bw() + 
  ylim(c(0,1)) + theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

#### Race

Since we are counting Hispanic as a race in this case, we will calculate Hispanic
as Hispanic Ethnicity Proportion * White Race Proportion, and adjust White Race 
Proportion to be Not Hispanic Ethnicity Proportion * White Race Proportion.

```{r}
Props %>%
  select(Descriptors$Description[27:34]) %>%
  mutate(
    Hispanic = White * Hispanic,
    White = White * `Not Hispanic`
  ) %>%
  select(-`Not Hispanic`) %>%
  pivot_longer(Descriptors$Description[c(27:32, 34)]) %>%
  rename(Race = name, Proportion = value) %>%
  ggplot(aes(x = Race, y = Proportion, fill = Race)) + geom_boxplot() + theme_bw() + 
  ylim(c(0,1)) + theme(legend.position = "none") 
```

```{r}
pairs(Props %>% select(Descriptors$Description[27:34]))
```

#### Education

```{r}
Props %>%
  select(Descriptors$Description[35:38]) %>% 
  pivot_longer(Descriptors$Description[35:38]) %>%
  rename(Education = name, Proportion = value) %>%
  mutate(Education = factor(Education, levels = Descriptors$Description[35:38])) %>%
  ggplot(aes(x = Education, y = Proportion, fill = Education)) + geom_boxplot() + theme_bw() + 
  ylim(c(0,1)) + theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
pairs(Props %>% select(Descriptors$Description[35:38]))
```

#### Sex

```{r}
Props %>%
  select(Descriptors$Description[39:40]) %>% 
  pivot_longer(Descriptors$Description[39:40]) %>%
  rename(Sex = name, Proportion = value) %>%
  mutate(Sex = factor(Sex, levels = Descriptors$Description[39:40])) %>%
  ggplot(aes(x = Sex, y = Proportion, fill = Sex)) + geom_boxplot() + theme_bw() + 
  ylim(c(0,1)) + theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Outlier Detection 

```{r}
# Run PCA on log adjusted data 
PCA <- prcomp(log10(RAC_County[,3:42] + 1e-10), scale = T)
PCA_Res <- data.table("PC1" = PCA$x[,1], "PC2" = PCA$x[,2], "FIPS" = RAC_County$FIPS, "Total" = log10(RAC_County$c000)) %>%
  merge(
    unique(RAC[,c("cty", "ctyname")]) %>% 
      mutate(ctyname = lapply(ctyname, function(x) {
        strsplit(x, ",") %>% unlist() %>% tail(1)}) %>% unlist()) %>% 
      rename(FIPS = cty, State = ctyname), 
   by = "FIPS") 

PCA_Res%>%
  ggplot(aes(x = PC1, y = PC2, color = Total, label = FIPS)) + geom_point() + xlab("PC1 (83.93%)") + 
    ylab("PC2 (6.83%)") + theme_bw()
```

There may be 5 outliers, which are small counties with small to non-existant minority
populations and without certain industries.

```{r}
PotOut <- PCA_Res %>% filter(PC2 < -5) %>% select(FIPS) %>% unlist()

datatable(RAC_County %>% filter(FIPS %in% PotOut), options = list(scrollX = T))
```



