---
title: "Unequal Commute Data Analysis"
author: "Anastasiya Prymolenna"
date: "12/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(geojsonR)
library(geojsonio)
library(broom)
library(ggplot2)
library(geosphere)
library(maps)
#library(OpenStreetMap)
library(dplyr)
library(viridis)
library(reshape2)
library(leaflet)
```


```{r David}

# covid <- "../og_data/job_access/job_loss_by_tract.geojson"
# covid_data <-  geojson_read(covid,  what = "sp")
```

Notes:

Java required for OpenStreetMap, which doesnt require Google API to use maps for spatial plotting, unlike ggmap.

Questions:
levels the data is observed at
years the data is collected
normalization explained. What is it normalized to
spacial domain

External Data Sets ideas:

1. Child care accessibility
2. Lisa's education data set
3. BLS career opportunities within the GEOIDs 


## Data input

The data behind the Urban Institute's Unequal Commute feature, published on October 6, 2020.

Takeaways from the report:
 
 1. Access to jobs varies by neighborhood
 2. A lack of public transit access hinders transit dependent workers’ ability to commute
 3. People of color have less access to safe and affordable transportation
 4. Better data are needed to help increase transportation equity and access to opportunity



It is restricted to only the access to job opportunity in the Lansing, Michigan; Seattle, Washington;
Baltimore, Maryland; and Nashville, Tennessee MSAs.
Each census block represented by a GEOID is a subdivision of a census tract that generally contains between 600 and 3,000 people.

Job access is calculcated based on a gravity model of the combined auto and transit access for low wage workers in a block group.
This is defined by the number of jobs in an area multiplied by the impidence function of travel time divided by the sum of
the total population of commuters and the impidence function of travel time for all of them. 

```{r input files}
file_js_path <- "../../og_data/job_access/job_access_gap.json"
job_access_csv <- "../../og_data/job_access/job_access_gap.csv"
job_access_gap_dict <- "../../og_data/job_access/job_access_gap_dict.csv"
job_access_equity <- "../../og_data/job_access/job_access_gap_equity.csv"
job_access_gap_equity_dict <- "../../og_data/job_access/job_access_gap_equity_dict.csv"


file_js = FROM_GeoJson(url_file_string = file_js_path)
job_access <- read.csv(job_access_csv)
job_access_gap <- read.csv(job_access_equity)
definitions <- read.csv(job_access_gap_dict)
definitions
```


Leading Questions:
How is the overall job access related to the block group workers commute from?
Is overall job access more correlated with day time versus night time workers?
How has COVID impacted the block regions and job access in these four metropolitan areas.
Do people of color have equal access to job opportunities?

```{r centroid function}
cntrd <- function(x) {
  as.vector(centroid(as.matrix(x[,c("long", "lat")])))
}
```

```{r view data}
#Load data as a spacial object
jadf <- geojson_read(file_js_path,  what = "sp")

#subset specific region
jadf_WA <- jadf[ jadf@data$MSA == "Seattle" , ]
jadf_WA_formatted <- tidy(jadf_WA)

jadf_MI <- jadf[ jadf@data$MSA == "Lansing" , ]

jadf_MD <- jadf[ jadf@data$MSA == "Baltimore" , ]

jadf_TN <- jadf[ jadf@data$MSA == "Nashville" , ]


#get map info
center_input <- cntrd(jadf_WA_formatted)
#map <- openmap(c(center_input[1]+0.7, center_input[2]-0.7), c(center_input[1]-0.7, center_input[2]+0.7),
#               type="osm", zoom=FALSE)

# Plot it
ggplot() +
  geom_polygon(data = jadf_WA_formatted, aes( x = long, y = lat, group = group)) +
  theme_void() +
  coord_map()

pal <- colorNumeric("viridis", NULL)

```

```{r, plot spacial data}
leaflet(jadf_WA) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
    fillColor = ~pal(OverallJobAccess),
    label = ~paste0(GEOID, ": ", formatC(OverallJobAccess, big.mark = ","))) %>%
  addLegend(pal = pal, values = ~OverallJobAccess, opacity = 1.0)

leaflet(jadf_MI) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
    fillColor = ~pal(OverallJobAccess),
    label = ~paste0(GEOID, ": ", formatC(OverallJobAccess, big.mark = ","))) %>%
  addLegend(pal = pal, values = ~OverallJobAccess, opacity = 1.0)

leaflet(jadf_MD) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
    fillColor = ~pal(OverallJobAccess),
    label = ~paste0(GEOID, ": ", formatC(OverallJobAccess, big.mark = ","))) %>%
  addLegend(pal = pal, values = ~OverallJobAccess, opacity = 1.0)

leaflet(jadf_TN) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
    fillColor = ~pal(OverallJobAccess),
    label = ~paste0(GEOID, ": ", formatC(OverallJobAccess, big.mark = ","))) %>%
  addLegend(pal = pal, values = ~OverallJobAccess, opacity = 1.0)

```


Time of Day Comparison

```{r}
leaflet(jadf_MD) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
    fillColor = ~pal(Access30Transit),
    label = ~paste0(GEOID, ": ", formatC(Access30Transit, big.mark = ","))) %>%
  addLegend(pal = pal, values = ~Access30Transit, opacity = 1.0)

leaflet(jadf_MD) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
    fillColor = ~pal(Access30Transit_ts),
    label = ~paste0(GEOID, ": ", formatC(Access30Transit_ts, big.mark = ","))) %>%
  addLegend(pal = pal, values = ~Access30Transit_ts, opacity = 1.0)
```


## Explore Data


```{r}
job_access_groups <- job_access %>% group_by(MSA)

ggplot()+
  geom_boxplot(data=job_access_groups, aes(x= MSA, y=OverallJobAccess)) +
  ggtitle("Normalized measure of overall access to jobs for low wage job seekers")

ggplot()+
  geom_point(data=job_access_groups, aes( x=OverallJobAccess,y= spatialmismatch, colour=Access30Transit)) +
  facet_wrap(~MSA)+
  ggtitle("acessibility of jobs compared to spacial mismatch between job seekers and jobs")

ggplot()+
  geom_point(data=job_access_groups, aes( x=lowwagelaborforce,y= spatialmismatch, colour =OverallJobAccess)) +
  facet_wrap(~MSA)+
  ggtitle("acessibility of jobs per block group from LODES data")

ggplot()+
  geom_point(data=job_access_groups, aes( x=Access30Transit,y= spatialmismatch,  colour =OverallJobAccess)) +
  facet_wrap(~MSA)+
  ggtitle("time to get to work on spacial missmatch")

ggplot(job_access, aes(lowwagelaborforce, OverallJobAccess, labels=MSA, colour = MSA)) + scale_fill_viridis() +
  geom_point()+
  ggtitle("Measure of how long different groups of the labor force take to get to work")


ggplot()+
  geom_boxplot(data=job_access_groups, aes(x= MSA, y=Access30Transit))
```


```{r}
job_access_gap_melt <- melt(job_access_gap)
job_melt_subset <- subset(job_access_gap_melt, job_access_gap_melt$variable!="total_pop")

ggplot(job_melt_subset, aes(fill=variable, y=value, x=type))+
  geom_bar(position="dodge", stat="identity")+
  facet_wrap(~cbsa_name)+
  ggtitle("Demographic information of each city")

```




```{r}

```







